---
title: "Application Examples"
author: "Aimie Doriath Döhler"
format:
  html:
    toc: true
    toc-depth: 3
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

In this vignette, we:

1.  Show application examples of the aperture shape as a new
    development-oriented framework for ammonite morphometrics:

    -   **Intrageneric variation**

    -   **Ontogenetic trajectories**

    -   **Disparity**

2.  Compare the results of 3D aperture shape to classic descriptors (rib
    shapes in 2D and whorl section shape) in term of **taxa
    discrimination**

Detailed methodological descriptions, statistical rationale, and
biological interpretations provided in the paper are not repeated here.
The goal is not to interpret biological results, but to **demonstrate
how those results were obtain** and therefore ensuring **computational
transparency and full reproducibility** of the analyses.

::: callout-note
The workflow of data acquisition, preprocessing, and missing data
reconstruction are detailed in the vignette [Getting started with
AmmoniTools](Getting_started_with_AmmoniTools.html), while evaluation of
the potential methodological biases are presented in [Test
biases](Test_biases.html) vignette.
:::

# Data setup and preprocessing

For this application examples, we used the example data set provided
with the package. Every curve is resampled with 25 landmarks.
Symmetrisation were computed using exponential weighting. Reconstruction
of missing data using TPS-based interperistome reconstruction were
compute using a hierarchical structure (Specimen -\> Species -\> Genus)
and a exponential weighting of complete peristome in the reconstruct
peristome.

(See [*Getting started with
AmmoniTools*](Getting_started_with_AmmoniTools.html) for more
information of the workflow of data acquisition, preprocessing, and
missing data reconstruction.)

```{r setup_and_preprocessing, message = FALSE, warning=FALSE, results='hide'}
library(AmmoniTools)
library(ggplot2)
library(dplyr)
library(Morpho)

# Load dataset
data("example_spec_tab", package = "AmmoniTools")
data("example_data", package = "AmmoniTools")

# Resample curve with 25 landmarks for each curve type
resampled_data <- resample_curves(example_data, n_landmarks = list(R = 25, L = 25, U = 25, V = 25))

# Reorder curves
reordered_data <- reorder_landmarks(resampled_data)

# Symmetrize curves
data_sym <- symmetrize_sides(reordered_data,
                             weighting = "exp" # example with exponential weighting
                             )

# merged R and L sides as one object
merged_data <- merge_sides(data_sym)

# reconstruct missing landmarks
reconstructed_data <- reconstruct_missingldm(
  merged_data,
  metadata = example_spec_tab,
  hierarchy = list("Specimen", "Species", "Genus"),
  weighting_function = "gaussian",
  max_donors = 5, verbose = TRUE)
```

# 1. Intrageneric variations

This section reproduces the analyses to obtain [**figure
10**]{.underline} in the manuscript.

Here, we focused on population of *Pleydellia* specimens.

```{r disparity_sample}
# Subsetting the dataset
data_Pleydellia <- subset_data(reconstructed_data, example_spec_tab,
                                Genus = c("Pleydellia"))
```

A morphospace representing the variation of *Pleydellia* peristomes is
obtain by computing a General Procruste Alignment \[Proc_peristome\]
followed by a Principal Component Analysis \[compile_pca\]. The
morphospace can be visualised using
[ggplot2::ggplot()](https://www.rdocumentation.org/packages/ggplot2/versions/4.0.1/topics/ggplot).

```{r disparity_gpa_pca, message = FALSE, warning=FALSE, results='hide'}
gpa_Pleydellia <- Proc_peristome(data_Pleydellia, slide = TRUE, fixed_ldm = c(1, 25, 49))
#plot(gpa)
res <- compile_pca(gpa_Pleydellia, data_Pleydellia, example_spec_tab, filter_pc01 = TRUE)

pca_scores_Pleydellia <- res$pca_scores
variance_Pleydellia <- res$variance
pca_Pleydellia <- res$pca
```

```{r disparity_gpa_pca2}
var_explained <- round(100 * (pca_Pleydellia$sdev^2 / sum(pca_Pleydellia$sdev^2)), 1)

var_pc1 <- var_explained[1]  # PC1
var_pc2 <- var_explained[2]  # PC2

# PC1 VS PC2
p <- ggplot(pca_scores_Pleydellia, aes(x = Comp1, y = Comp2)) +
  geom_point(size = pca_scores_Pleydellia$Csize/30, alpha = 1) +
  theme_minimal() +
  coord_fixed()+
  labs(title = "Morphospace of Pleydellia peristomes.", 
       x = paste0("PC1 (", var_pc1, "%)"),
       y = paste0("PC2 (", var_pc2, "%)"))
#ggsave("Pleydellia_morphospace.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```

Theoritical minimal (in blue), mean (in black) and maximal (in red)
shapes of Principal Components can be visualized like this:

```{r disparity_shapes}
plot_pca_shapes3d(gpa_Pleydellia, pca_Pleydellia, pcs = 1:2)
```

The first row represent PC1 shape deformation whereas second row
represent PC2's.

# 2. Ontogenetic trajectories

This section reproduces the analyses to obtain [**figure
11**]{.underline} in the manuscript.

**Ontogenetic trajectories**, i.e. shape variations through ontogeny,
can be obtain by ploting PC against Csize.

```{r onto_traj, message=FALSE, warning=FALSE, results='asis', fig.align='center', fig.width=8, fig.height=5}
cac_res <- CAC(pca_scores_Pleydellia[,1:8], pca_scores_Pleydellia$Csize, log = FALSE)

cac_df <- data.frame(
  CACscore  = cac_res$CACscores,
  Csize     = pca_scores_Pleydellia$Csize,
  Specimen  = pca_scores_Pleydellia$Specimen,
  Genus     = pca_scores_Pleydellia$Genus
)

# Compute mean centroid size per specimen
specimen_order <- cac_df %>%
  group_by(Specimen) %>%
  summarise(mean_Csize = mean(Csize, na.rm = TRUE)) %>%
  arrange(mean_Csize) %>%
  pull(Specimen)

# Reorder Specimen factor
cac_df <- cac_df %>%
  mutate(Specimen = factor(Specimen, levels = specimen_order))

# Generate a custom blue → black → red palette
n_specimens <- length(unique(cac_df$Specimen))
custom_palette <- colorRampPalette(c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))(n_specimens)

# Plot
p <- ggplot(cac_df, aes(x = Csize, y = CACscore, group = Specimen, color = as.factor(Specimen))) +
  geom_line(linewidth = 0.8, alpha = 0.7) +
  geom_point(size =2, alpha = 0.7, shape = 16, stroke = 0, alpha = 0.8) +
  scale_color_manual(values = custom_palette) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Pleydellia ontogenetic trajectories (CAC)",
    x = "Centroid size",
    y = "CAC score",
    color = "Specimen"
  )
# ggsave("Pleydellia_ontotraj.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p

plot_cac_shapes3d(gpa_Pleydellia, pca_Pleydellia, cac_res)
```

# 3. Disparity

This section reproduces the analyses to obtain [**figure
12**]{.underline} in the manuscript.

The dataset was subsampled by selecting specimens of *Grammoceras
penestriatulum*, *Harpoceras* *subplanatum* and *Leioceras opalinum*:

```{r discr_sampling, message=FALSE, warning=FALSE, results='hide'}
data_evol <- subset_data(reconstructed_data, example_spec_tab,
                                Species = c("aalensis",
                                            "penestriatulum",
                                            "subplanatum"
                                            ))
```

Using a similar method as section 2, a morphospace visualization is
obtained as well as the deformation shapes along the PC axis:

```{r disparity_evol, message=FALSE, warning=FALSE, results='hide'}
gpa <- Proc_peristome(data_evol, slide = TRUE, fixed_ldm = c(1, 25, 49))
plot(gpa)
res <- compile_pca(gpa, data_evol, example_spec_tab)

pca_scores <- res$pca_scores
variance <- res$variance
pca <- res$pca
```

```{r disparity_evol_PC}
var_explained <- round(100 * (pca$sdev^2 / sum(pca$sdev^2)), 1)

var_pc1 <- var_explained[1]  # PC1
var_pc2 <- var_explained[2]  # PC2

# PC1 VS PC2
p <- ggplot(pca_scores, aes(x = Comp1, y = Comp2, color = as.factor(Genus))) +
    geom_point(size =2, alpha = 0.4, shape = 16, stroke = 0, alpha = 0.8) +
    theme_minimal() + 
    geom_polygon(
      data = pca_scores %>% group_by(Genus) %>% 
        slice(chull(Comp1, Comp2)),  # convex hull
      aes(fill = Genus),
      alpha = 0.2,
      linewidth = 1,
      show.legend = FALSE 
    ) +
   coord_fixed()+
   labs(title = "Morphospace at evolutionary scale", 
        x = paste0("PC1 (", var_pc1, "%)"),
        y = paste0("PC2 (", var_pc2, "%)"),
        color = "Genus")
#ggsave("evol_morphospace.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
# Plot shape deformation along axis
plot_pca_shapes3d(gpa, pca, pcs = 1:2)
```

# 4. Taxa discrimination

This section reproduces the analyses used to obtain [**Figure
13**]{.underline} in the manuscript.

All analyses are performed on the same dataset (see Section 4),
including five specimens of each species: *G. penestriatulum*, *H.
subplanatum* and *L. opalinum*.

The goal is to compare the taxonomic discriminant power of full 3D
aperture shapes, 2D rib outlines (lateral projection) and 2D whorl
section outlines (frontal projection).

To ensure comparability, all datasets follow the same analytical
pipeline:

-   extraction of Procrustes-aligned coordinates,

-   optimization of PCA dimensionality using LDA leave-one-out
    cross-validation,

-   computation and visualization of discriminant morphospaces.

## Extraction of Procrustes-aligned coordinates

### 3D aperture shapes

```{r taxo_3Dap, message=FALSE, warning=FALSE, results='hide'}
data_evol <- subset_data(
  reconstructed_data, example_spec_tab,
  Species = c("aalensis", "penestriatulum", "subplanatum")
)

gpa_3D <- Proc_peristome(
  data_evol,
  slide = TRUE,
  fixed_ldm = c(1, 25, 49)
)

res_3D <- compile_pca(gpa_3D, data_evol, example_spec_tab)

pca_scores_3D <- res_3D$pca_scores
```

### 2D rib shapes (lateral projection)

```{r taxo_2Dribs, message=FALSE, warning=FALSE, results='hide'}
twod_ribs <- project_on_plane(
  data = data_evol,
  k = 5,
  mode = "2D_lateral",
  plot = FALSE
)

gpa_ribs <- Projection_2D_gpa(
  twod_ribs,
  slide = TRUE,
  fixed_ldm = c(1, 25)
)

res_ribs <- compile_pca(gpa_ribs$gpa, data_evol, example_spec_tab)

pca_scores_ribs <- res_ribs$pca_scores
```

### 2D whorl section shapes (frontal projection)

```{r taxo_2Dsection, message=FALSE, warning=FALSE, results='hide'}
whorl_section <- project_on_plane(
  data = data_evol,
  mode = "2D_frontal"
)

gpa_ws <- Projection_2D_gpa(
  whorl_section,
  slide = TRUE,
  fixed_ldm = c(1, 25, 49)
)

res_ws <- compile_pca(gpa_ws$gpa, data_evol, example_spec_tab)

pca_scores_ws <- res_ws$pca_scores
```

## Optimization of PCA dimensionality using LDA cross-validation

To objectively determine the number of principal components to retain,
we computed leave-one-out LDA classification accuracy as a function of
PCA dimensionality for each dataset.

**Compute accuracies for the three datasets**

```{r}
max_pc = 35
acc_3D   <- compute_lda_accuracy(pca_scores_3D,   max_pc, group_var = "Genus")
acc_ribs <- compute_lda_accuracy(pca_scores_ribs, max_pc, group_var = "Genus")
acc_ws   <- compute_lda_accuracy(pca_scores_ws,   max_pc, group_var = "Genus")

acc_3D$dataset   <- "3D apertures"
acc_ribs$dataset <- "2D ribs"
acc_ws$dataset   <- "2D whorl section"

acc_all <- rbind(acc_3D, acc_ribs, acc_ws)
```

**Combined optimization plot (Figure S1)**

```{r}
p <- ggplot(acc_all, aes(x = n_pc, y = accuracy, color = dataset)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  geom_point(size = 1.5) +
  theme_minimal(base_size = 14) +
  geom_vline(xintercept = 20, linetype = "dashed", linewidth = 0.8) +
  scale_color_manual(
    values = c(
      "3D apertures"   = "#2a9d8f",
      "2D ribs" = "#f4a261",
      "2D whorl section"     = "#e76f51"
    )
  ) +
  labs(
    x = "Number of retained PCs",
    y = "Classification accuracy",
    color = "Dataset",
    title = "Optimization of PCA dimensionality for LDA"
  )
#ggsave("PC_optimization.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```

👉 Based on this analysis, we retained n_pc = 20 for all datasets,
providing a compromise between classification performance and
comparability.

## Discriminant morphospaces using optimized PCA dimensionality

Using the optimized number of PCs (n_pc = 20), we computed LDA models
and visualized the resulting discriminant morphospaces.

### 2D ribs (Fig. 13A)

```{r rib_morphospace}
lda_df_ribs <- run_lda(pca_scores_ribs)
p <- ggplot(lda_df_ribs, aes(LD1, LD2, color = Genus)) +
  geom_point(size =2, alpha = 0.4, shape = 16, stroke = 0, alpha = 0.8) +
  geom_polygon(
    data = lda_df_ribs %>% group_by(Genus) %>% slice(chull(LD1, LD2)),
    aes(fill = Genus),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  coord_equal() +
  theme_minimal(base_size = 14) +
  labs(title = "2D rib shapes")
# ggsave("LDA_2Drib.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```

### 2D whorl sections (Fig. 13B)

```{r section_morphospace}
lda_df_ws <- run_lda(pca_scores_ws)

p <- ggplot(lda_df_ws, aes(LD1, LD2, color = Genus)) +
  geom_point(size =2, alpha = 0.4, shape = 16, stroke = 0, alpha = 0.8) +
  geom_polygon(
    data = lda_df_ws %>% group_by(Genus) %>% slice(chull(LD1, LD2)),
    aes(fill = Genus),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  coord_equal() +
  theme_minimal(base_size = 14) +
  labs(title = "2D whorl section shapes")
# ggsave("LDA_2Dwhorlsection.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```


### 3D apertures (Fig. 13C)

```{r ap_morphospace}
lda_df_ap <- run_lda(pca_scores_3D)

p <- ggplot(lda_df_ap, aes(LD1, LD2, color = Genus)) +
  geom_point(size =2, alpha = 0.4, shape = 16, stroke = 0, alpha = 0.8) +
  geom_polygon(
    data = lda_df_ap %>% group_by(Genus) %>% slice(chull(LD1, LD2)),
    aes(fill = Genus),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  coord_equal() +
  theme_minimal(base_size = 14) +
  labs(title = "3D aperture shapes")
# ggsave("LDA_3Daperture.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```
