---
title: "Bias evaluation"
author: "Aimie Doriath Döhler"
format:
  html:
    toc: true
    toc-depth: 3
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  warning = FALSE,
  message = FALSE
)
devtools::load_all()
library(AmmoniTools)
```

# Introduction

This vignette documents the **evaluation of methodological biases** associated with the acquisition of ammonite 3D models using surface scanning and photogrammetry, the landmarking of apertures using [3D Slicer](https://www.slicer.org/) software, and the reconstruction of ammonite peristomes using **`AmmoniTools`**. The goal is not to interpret biological results, but to **demonstrate how biases was quantified** and to ensure **computational transparency and full reproducibility** of the analyses.

Detailed methodological descriptions, statistical rationale, and biological interpretations provided in the paper are not repeated here. However, this section present supplementary analyses and figures not included in the manuscript, which are intended to further document the behaviour and robustness of the methods.

::: callout-note
Data acquisition, preprocessing, and missing data reconstruction are detailed in [Getting started with AmmoniTools](Getting_started_with_AmmoniTools.html), while illustrative biological applications are presented in [Applications examples](Applications_examples.html).
:::

# Imaging bias

This section reproduces the analyses and figures used to quantify imaging-related bias ([**Figure 5**]{.underline} in the manuscript).

```{r imaging_bias}
library(readxl)
library(tidyr)
library(ggplot2)

# Load data
file_path <- rprojroot::find_package_root_file("inst", "extdata", "quantif_deviation.xlsx")

Geomagic_data <- read_excel(file_path) # Tab with specimen information

Geomagic_data <- Geomagic_data[1:6, ]

# Selection of column
Geomagic_long <- Geomagic_data %>%
  pivot_longer(
    cols = c(`Mean distance (mm)`,
             `Ecart type (mm)`,
             `Estimation RMS (mm)`),
    names_to = "Metric",
    values_to = "Value"
  )
```

```{r imaging_bias2}
p <- ggplot(Geomagic_long, aes(x = Metric, y = Value, fill = imaging_technics)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA, position = position_dodge(width=0.9)) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Imaging bias test results",
       y = "Metric value (mm)",
       fill = "Imaging technique",
       color = "Imaging technique")
# ggsave("Imaging_Bias.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```

# Manual landmarking bias

This section reproduces the analyses assessing variability introduced by manual landmark placement ([**Figure 7**]{.underline} in the manuscript).

```{r user_bias_data, message=FALSE, warning=FALSE, results='hide'}
library(geomorph)
library(dplyr)

# Load data (from extdata file in AmmoniTools package files)
file_path <- rprojroot::find_package_root_file("inst", "extdata", "spec_tab_userbias.xlsx")
spec_tab <- read_excel(file_path) # Tab with specimen information
folder <- rprojroot::find_package_root_file("inst", "extdata")

# Extract ldm
raw_data <- extract_landmarks(folder, spec_tab,
                              expected_ldm = list(R = 200, L = 200, U = 100, V = 100))
# no missing ldm
raw_data$curve_info$missing_ldm <- rep(list(integer(0)), nrow(raw_data$curve_info))

# Reorder ldm + merge sides
reordered_data <- reorder_landmarks(raw_data)
data <- merge_sides(reordered_data)

# GPA
n_ldm <- nrow(data$landmarks[[1]][[1]]) # Nb of landmarks
gpa <- Proc_peristome(data, slide = TRUE,
                      fixed_ldm = c(1, (n_ldm+1)/4, (n_ldm+1)/2, 3*(n_ldm+1)/4-1, n_ldm))

# PCA
pca <- gm.prcomp(gpa$coords)

# PCA scores
pca_scores <- as.data.frame(pca$x)

stages <- integer()
for (specimen_id in names(data$landmarks)) {
  # get only apertures (numbers)
  aperture_names <- grep("^\\d+$", names(data$landmarks[[specimen_id]]), value = TRUE)
  aperture_names <- aperture_names[order(as.numeric(aperture_names))]  # numeric order

  # assign stage numbers
  stages <- c(stages, seq_along(aperture_names))
}

# Add to PCA scores
pca_scores$Stage <- stages

specimen = character() # get specimen id
for (specimen_id in names(data$landmarks)) {
  aperture_names <- grep("^\\d+$", names(data$landmarks[[specimen_id]]), value = TRUE)
  for (i in seq_along(aperture_names)) {
    specimen <- rbind(specimen, specimen_id)
  }
}
pca_scores$Specimen <- as.numeric(specimen[, 1])
pca_scores$Csize <- gpa$Csize

# Fused on common id (=Specimen)
pca_scores <- pca_scores %>%
  left_join(spec_tab, by = "Specimen")
```

Here is the code to plot the PCA morphospace resulting of the manual landmarking repeatability test ([**figure 7A**]{.underline} in the manuscript):

```{r user_bias_A}
hulls <- pca_scores %>%
  group_by(Stage) %>%
  slice(chull(Comp1, Comp2))
p <- ggplot(pca_scores, aes(x = Comp1, y = Comp2)) +
  geom_polygon(data = hulls, aes(fill = as.factor(Stage), group = Stage),
               alpha = 0.3, color = NA) +
  geom_point(aes(color = as.factor(Stage)), size = 2) +
  theme_minimal() +
  coord_fixed()+
  labs(title = "PCA Morphospace resulting from the manual \nlandmarking repeatability test",
    x = "PC1", y = "PC2",
       color = "Replicats",
       fill = "Replicats")
# ggsave("User_bias_PCA_plot.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```

Here is the code to plot the PCA morphospace resulting of the manual landmarking repeatability test ([**Figure 7D**]{.underline} in the manuscript):

```{r user_bias_D}
p <- ggplot(pca_scores, aes(x = Csize, y = Comp1, group=Specimen, color = as.factor(Stage))) +
  geom_line(color="grey",linewidth = 0.8, alpha = 0.7) +
  geom_point(size=2) +
  theme_minimal() +
  labs(title ="Ontogenetic trajectory (PC1 vs. Centroid size) resulting \n
       from the manual landmarking repeatability test, with replicates \n
       connected by grey lines.",
  x = "Csize", y = "PC1",
       color = "Replicats")
# ggsave("User_bias_ontotraj.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p
```

Here is the code to plot the mean mean Procrustes residuals value par landmarks mapped onto the mean peristome shape ([**Figure 7C**]{.underline}):

```{r user_bias_C}
# Procruste distance by stage
coords <- gpa$coords
stage_vec <- pca_scores$Stage

procrustes_distances <- lapply(unique(stage_vec), function(st) {
  idx <- which(stage_vec == st)
  if (length(idx) > 1) {
    n <- length(idx)
    d <- matrix(0, n, n)
    for (i in 1:(n-1)) {
      for (j in (i+1):n) {
        d[i,j] <- sqrt(sum((coords[,,idx[i]] - coords[,,idx[j]])^2))
        d[j,i] <- d[i,j]
      }
    }
    return(list(stage = st, distances = d, specimens = pca_scores$Specimen[idx]))
  } else {
    return(list(stage = st, distances = NA, specimens = pca_scores$Specimen[idx]))
  }
})

coords <- gpa$coords     # array p x k x n (landmarks x dims x specimens)
stage_vec <- pca_scores$Stage
p <- dim(coords)[1]      # number of landmarks
k <- dim(coords)[2]      # dimensions

# storage
mean_distances_by_ldm <- list()

for (st in unique(stage_vec)) {
  idx <- which(stage_vec == st)

  if (length(idx) > 1) {
    # tab to store distances by landmarks
    dist_mat <- matrix(NA, nrow = length(idx), ncol = p)
    all_pairs <- combn(idx, 2)
    dists_ldm <- matrix(0, nrow = ncol(all_pairs), ncol = p)

    for (c in 1:ncol(all_pairs)) {
      i <- all_pairs[1, c]
      j <- all_pairs[2, c]
      diff <- coords[,,i] - coords[,,j]  # p x k
      dists_ldm[c, ] <- sqrt(rowSums(diff^2)) # vector p
    }

    # mean distance for each landmark
    mean_distances_by_ldm[[as.character(st)]] <- colMeans(dists_ldm)
  } else {
    mean_distances_by_ldm[[as.character(st)]] <- rep(NA, p)
  }
}

# convert into data.frame long for ggplot
df_ldm <- do.call(rbind, mean_distances_by_ldm)
df_ldm <- as.data.frame(df_ldm)
df_ldm$Stage <- rownames(df_ldm)

df_ldm_long <- df_ldm %>%
  pivot_longer(cols = -Stage, names_to = "Landmark", values_to = "Mean_Procrustes_Distance") %>%
  mutate(Stage = as.numeric(Stage), Landmark = as.numeric(gsub("V", "", Landmark)))

# Summary by landmark
dist_summary <- df_ldm_long %>%
  group_by(Landmark) %>%
  summarise(
    Mean = mean(Mean_Procrustes_Distance, na.rm = TRUE),
    Min = min(Mean_Procrustes_Distance, na.rm = TRUE),
    Max = max(Mean_Procrustes_Distance, na.rm = TRUE),
    Q1 = quantile(Mean_Procrustes_Distance, 0.25, na.rm = TRUE),
    Q3 = quantile(Mean_Procrustes_Distance, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

p <- ggplot(dist_summary, aes(x = Landmark)) +
  geom_ribbon(aes(ymin = Min, ymax = Max), fill = "gray80", alpha = 0.3) +
  geom_ribbon(aes(ymin = Q1, ymax = Q3), fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = Mean), color = "steelblue", size = 1.2) +
  labs(
    title = "Supplementary figure: Mean Procruste distance \n
    by landmarks resulting from the manual landmarking \n
    repeatability test",
    x = "Landmarks index",
    y = "Procruste distance"
  ) +
  theme_minimal(base_size = 14)
# ggsave("User_bias_distbyldm", plot = p, width = 8, height = 6, units = "in", device = "pdf")
p

# Procruste distance on mean gpa coords
# Mean by landmark
mean_shape <- apply(coords, c(1, 2), mean)  # dim: n_landmarks x 3

# Variance
var_ldm <- dist_summary %>% arrange(Landmark) %>% pull(Mean)

palette_fun <- colorRampPalette(c("blue", "green", "yellow", "red"))
n_colors <- 100
colors <- palette_fun(n_colors)

# Normalise variance
var_norm <- (var_ldm - min(var_ldm)) / (max(var_ldm) - min(var_ldm))
color_indices <- round(var_norm * (n_colors - 1)) + 1
ldm_colors <- colors[color_indices]

# 3D plot
rgl::open3d(silent = TRUE)
spheres3d(mean_shape, radius = 0.003, color = ldm_colors)
lines3d(mean_shape, color = "gray")

# rgl.postscript("User_bias_distanceldm.pdf", fmt = "png")
# rgl.snapshot("User_bias_distanceldm.png", fmt = "png")

## A random peristome
spheres3d(mean_shape, radius = 0.001, color = "indianred2")
#rgl.snapshot("random_peristome.png", fmt = "png")
rgl::rglwidget(width = 600, height = 600)
```

Here is the code used for the resampling test ([**Figure 7B**]{.underline} in the manuscript):

```{r user_bias_B}
# var_intra : intrastage variance vector
# var_inter : interstage variance vector
# Data
df <- pca_scores
pc_cols <- grep("^Comp", names(df), value = TRUE)
k <- 3  # number of rows per sample

# Compute mean intra-stage variance
var_intra <- numeric()
for(stage in unique(df$Stage)){
  df_stage <- df[df$Stage == stage, ]
  n_stage <- nrow(df_stage)

  if(n_stage >= k){
    # All combinations of k rows within the stage
    combs <- combn(n_stage, k)
    for(i in 1:ncol(combs)){
      sampled <- df_stage[combs[, i], ]
      # Sum of variances across PCs
      var_intra <- c(var_intra, sum(sapply(pc_cols, function(pc) var(sampled[[pc]]))))
    }
  }
}
mean_varintra <- mean(var_intra)  # average intra-stage variance

# Compute inter-stage variance (all combinations globally)
n <- nrow(df)
combis <- combn(n, k)
Replicat_var <- numeric(ncol(combis))
count <- 0

for(i in 1:ncol(combis)){
  sampled_rows <- df[combis[, i], ]
  # Variance for each PC
  varianceintertot <- sapply(pc_cols, function(pc) var(sampled_rows[[pc]]))
  total_varianceinter <- sum(varianceintertot)
  Replicat_var[i] <- total_varianceinter

  # Count cases where mean intra-stage variance exceeds inter-stage variance
  if(mean_varintra > total_varianceinter){
    count <- count + 1
  }
}

# Compute p-value for comparison
p_value <- (count + 1) / (ncol(combis) + 1)
p_value

# Plot histogram of inter-stage variances
# pdf("inter_peristome_variance_hist.pdf", width = 7, height = 5)

hist(Replicat_var, breaks = 75,
     xlab = "Inter-peristome variance", ylab = "Frequency", col = "white", border = "black",
       main = "Distribution of inter-peristome variance",)
# Add line for mean intra-stage variance
abline(v = mean_varintra, col = "red", lwd = 2)
legend("topright", legend = "Mean intra-replicat variance", col = "red", lwd = 2)

# dev.off()

```

# Missing landmarks reconstruction bias

This section reproduces the analyses and figures used to quantify missing landmarks reconstruction bias ([**figures 8-9**]{.underline} in the manuscript). Some supplementary figures and information are provided here.

The analyses of this section are based on the example dataset provided with the package.

```{r setup, message=FALSE, warning=FALSE}
library(AmmoniTools)

# Load example dataset
data("example_data", package = "AmmoniTools")
raw_data <- example_data

data("example_spec_tab", package = "AmmoniTools")
spec_tab <- example_spec_tab

# Resample curves
resampled_data <- resample_curves(raw_data, n_landmarks = list(R = 25, L = 25, U = 25, V = 25))

# Reorder ldm
reordered_data <- reorder_landmarks(resampled_data)
```

# Test the accuracy of bilateral symmetry-based reconstruction

## Peristomes filtering

Apertures with both left and right sides preserved are filtered:

```{r symmetry}
filtered_data <- filter_curves(reordered_data, complete_only=FALSE)
```

If you want the information on the dataset:

``` r
# count number of curves by curve type
table(filtered_data$curve_info$Type)

# count number of specimen
cat("number of specimens included :",
    length(unique(filtered_data$curve_info$Specimen))) 
```

## Error distance computing

For each specimen, one side was artificially removed and then reconstructed, allowing direct comparison between original and reconstructed coordinates. The mean distance between corresponding landmarks (excluding originally missing landmarks) was computed and standardized by whorl height.

```{r error_dist_calcul, message=FALSE, warning=FALSE}
dist_df <- test_symmetrize_sides(filtered_data, weighting_methods = c("exp"), weight_power_values = 5)

library(tidyverse)
library(dplyr)
dist_summary_H <- dist_df %>% dplyr::select(Specimen, Curve, Mean_Distance_H)
```

```{r error_dist_calcul2}
x <- dist_summary_H$Mean_Distance_H

cat(
  "Mean :", mean(x, na.rm = TRUE), "mm\n",
  "SD   :", sd(x, na.rm = TRUE),   "mm\n",
  "Min  :", min(x, na.rm = TRUE),  "mm\n",
  "Max  :", max(x, na.rm = TRUE),  "mm\n"
)
```

```{r error_dist, fig.width=7, fig.height=5, fig.align='center'}
# Boxplot of distance error
#pdf("boxplot_mean_distance_H.pdf", width = 5, height = 5)
boxplot(
  dist_summary_H$Mean_Distance_H,
  main = "Distribution of Mean Distance\nstandardized by whorl height",
  ylab = "Mean Distance std by whorl height",
  col = "lightblue",
  border = "black",
  boxwex = 0.35,
  lwd = 0.8,
  outline = FALSE
)
#dev.off()
```

## 🆕 Detection of problematic peristomes *(supp. info.)*

A overview of the distribution of errors between specimen can be obtained. This can help detect the problematic peristomes and recheck if there is a problem in landmarking. A common mistake is the mismatch between left and right sides.

```{r pbt_spec, fig.width=7, fig.height=5}
# Compute automatic threshold
Q1 <- quantile(dist_summary_H$Mean_Distance_H, 0.25, na.rm = TRUE)
Q3 <- quantile(dist_summary_H$Mean_Distance_H, 0.75, na.rm = TRUE)
IQR_val <- Q3 - Q1
threshold_auto <- Q3 + 1.5 * IQR_val

cat("Automatic threshold =", round(threshold_auto, 3), "\n")

# Detect "problematic" peristomes 
dist_summary_H_high <- dist_summary_H %>%
  filter(Mean_Distance_H > threshold_auto) %>%
  arrange(desc(Mean_Distance_H))

# print(head(dist_summary_H_high))

library(knitr)
library(kableExtra)
knitr::kable(head(dist_summary_H_high, 5), caption = "Most problematic peristomes") %>%
  kable_styling(font_size = 9, bootstrap_options = c("condensed", "striped"))
```

➡️*In this example dataset, the 6th aperture of the specimen n°81 have the higher error value.*

## 🆕 Finding best weight power parameter *(supp. info.)*

`Weighting_power` is a decay parameter controlling the effect of linear or exponential weighting. Here, error distances at different weighting power values were computed to get the best value possible. This error values are compared to a no weighting mode (represented by th dashed line).

```{r weight_power, message=FALSE}
# Mean distance when no weighting (global symmetry plane)
dist_df_none <- test_symmetrize_sides(filtered_data,
                             weighting_methods = c("none"))
mean_none <- mean(dist_df_none$Mean_Distance_H)

# Mean distance when weighting (local symmetry plane)
dist_df <- test_symmetrize_sides(filtered_data,
                        weighting_methods = c("exp", "linear"),
                        weight_power_values = c(seq(1, 5, by = 1)))

# Compute mean per weight_power and weighting
dist_summary <- dist_df %>%
  group_by(weighting, weight_power) %>%
  summarise(
    Mean_Distance_H = mean(Mean_Distance_H, na.rm = TRUE),
    .groups = "drop"
  )

library(ggplot2)

# Plot with lines for each weighting method
p <- ggplot(dist_summary, aes(x = weight_power, y = Mean_Distance_H,
                         color = weighting, fill = weighting, group = weighting))+
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_hline(yintercept = mean_none, linetype = "dashed", color = "black") +
  theme_minimal(base_size = 14) +
  labs(
    x = "Weight power parameter",
    y = "Mean reconstruction error",
    title = "Supplementary figure: Effect of weight power \nparameter on reconstruction error by weighting method",
    color = "Weighting method",
    fill = "Weighting method"
  )
p
```

➡️*This plot shows that increasing λ results in higher reconstruction error for both exponential and linear weighting functions. Best weighting power value for linear and exponential weighting is respectively equal to 1 and 2. At those values, reconstruction error is comparable between exponential and linear weighting, and slightly higher for no weighting mode.*

## 🆕 Comparing weighting methods results *(supp. info.)*

Here is a comparison between weighting methods, using the best weightpower values for linear and exponential weighting.

```{r weighting_methods, fig.width=7, fig.height=5}
dist_df_linear <- test_symmetrize_sides (filtered_data,
                                weighting_methods = c("linear"),
                                weight_power_values = 1)
dist_df_exp <- test_symmetrize_sides (filtered_data,
                                weighting_methods = c("exp"),
                                weight_power_values = 5)

dist_df <- rbind(dist_df_none, dist_df_linear, dist_df_exp)

# Boxplot + violin plot Mean Distance standardized by whorl height
ggplot(dist_df, aes(x = weighting, y = Mean_Distance_H, fill = weighting)) +
  geom_violin(trim = TRUE, alpha = 0.6) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.8) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Supplementary figure: Comparison of weighting methods",
    y = "Mean Distance (H)",
    x = "Weighting method"
  )

# Compute mean distances per weighting method
mean_dist_by_weighting <- dist_df %>%
  group_by(weighting) %>%
  summarise(
    #Mean_Distance = mean(Mean_Distance, na.rm = TRUE),
    #Mean_Distance_CS = mean(Mean_Distance_CS, na.rm = TRUE),
    Mean_Distance_H = mean(Mean_Distance_H, na.rm = TRUE),
    .groups = "drop"
  )
print(mean_dist_by_weighting)
```

```{r anova_weighting_methods}
# Mean comparison test
shapiro.test(dist_df$Mean_Distance_H) # normality non respected

kruskal.test(Mean_Distance_H ~ weighting, data = dist_df) # "non parametric ANOVA"
```

➡️*Although exponential weighting yielded marginally smaller mean distances, differences between methods were not statistically significant. Therefore, no approach outperformed the others significantly. It should be noted, however, that this dataset only includes planispiral ammonites. For trochospiral ammonites, a local symmetry plane would likely be required. Consequently, exponential weighting with weight power = 5 was adopted as the default setting.*

## Shape-related effect ?

Here is the code used for the [**figure 8B**]{.underline} in the manuscript:

```{r sym_shape_effect1,  message = FALSE, warning=FALSE, results='hide'}
# Error values
dist_df <- test_symmetrize_sides (filtered_data,
                        weighting_methods = c("exp"),
                        weight_power_values = 5)

data_sym <- symmetrize_sides(filtered_data, weighting = "exp") # Symmetrize and reconstructed RL sides

merged_data <- merge_sides(data_sym) ## Merge R and L sides

# Reconstruct missing ventral and dorsal landmarks using TPS
reconstructed_data <- reconstruct_missingldm(merged_data,
                                             metadata = spec_tab,
                                             hierarchy = list("Species", "Genus"),
                                             weight_power = 3,
                                             max_donors = 5,
                                             weighting_function = "gaussian",
                                             verbose = FALSE)

gpa <- Proc_peristome(reconstructed_data,
                      slide = TRUE,
                      fixed_ldm = c(1, 25, 49))
pca <- gm.prcomp(gpa$coords)

pca_scores <- as.data.frame(pca$x[,])

pca_scores$Mean_Distance_H <- dist_df$Mean_Distance_H

# Percentage of variance explained
var_exp <- (pca$sdev^2) / sum(pca$sdev^2) * 100
pc1_lab <- paste0("PC1 (", round(var_exp[1], 1), "%)")
pc2_lab <- paste0("PC2 (", round(var_exp[2], 1), "%)")
```

```{r sym_shape_effect2}
## Morphospace
p <- ggplot(pca_scores, aes(x = Comp1, y = Comp2, color = Mean_Distance_H)) +
  geom_point(size = pca_scores$Mean_Distance_H*20, shape = 16, stroke = 0, alpha = 0.8) +
  scale_color_viridis_c() +
  scale_size(range = c(1, 6), guide = "none") +
  theme_minimal(base_size = 14) +
  labs(
    title = "PCA morphospace with peristomes colour-coded \naccording to symmetric reconstruction error.
",
    x = pc1_lab,
    y = pc2_lab,
    color = "Mean Distance / H"
  )+
  coord_equal()
p
# ggsave("Shape_effect_sym.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
```

## 🆕 Resume symmetrization effect *(supp. info.)*

You can observed the decrease of missing landmarks after symmetrization:

```{r missing_ldm_hist, message=FALSE, warning=FALSE}
# Symmetry ponderation
data_missingsym <- symmetrize_sides(reordered_data, weighting = "exp")
merged_data <- merge_sides(data_missingsym)

# Reconstruct missing ventral and dorsal landmarks using TPS
reconstructed_data <- reconstruct_missingldm(merged_data,
                                             metadata = spec_tab,
                                             hierarchy = list("Species", "Genus"),
                                             weight_power = 3,
                                             max_donors = 5,
                                             weighting_function = "gaussian",
                                             verbose = TRUE)

# Pourcent of reconstructed ldm before symmetrization
pourcent_missing_beforesym <- sapply(
  data_missingsym$curve_info$missing_ldm_updated,
  function(x) length(x) / nrow(data_missingsym$landmarks[[1]][[1]]) * 100
)
pourcent_missing_beforesym <- pourcent_missing_beforesym[pourcent_missing_beforesym != 0]

# Pourcent of reconstructed ldm after symmetrization
pourcent_missing_aftersym <- sapply(
  reconstructed_data$curve_info$missing_ldm_updated,
  function(x) length(x) / nrow(reconstructed_data$landmarks[[1]][[1]]) * 100
)
pourcent_missing_aftersym <- pourcent_missing_aftersym[pourcent_missing_aftersym != 0]

# A little bit of statistics
sum_before <- summary(pourcent_missing_beforesym)
sum_after <- summary(pourcent_missing_aftersym)
cat("Symmetrization reduces missing landmark on average from", sum_before[4], "to", sum_after[4], "pourcent")

# Comparing histogram
hist_range <- range(c(pourcent_missing_beforesym, pourcent_missing_aftersym), na.rm = TRUE)

hist(pourcent_missing_beforesym,
     breaks = 10,
     col = rgb(1, 0, 0, 0.5),   # rouge semi-transparent
     xlim = hist_range,
     xlab = "Percentage of missing landmarks",
     main = "Missing landmarks before reconstruction")

hist(pourcent_missing_aftersym,
     breaks = 10,
     col = rgb(0, 0, 1, 0.5),
     xlim = hist_range,
     xlab = "Percentage of missing landmarks",
     main = "Missing landmarks after reconstruction")
```

➡️*Symmetrization reduced the proportion of missing landmarks from an average of 12 % before symmetrization to 5 % afterward, demonstrating its effectiveness in improving data completeness prior to further reconstruction steps.*

# Test the accuracy of TPS-based interperistome reconstruction

When missing landmarks remain, i.e. missing on both sides, missing regions is reconstructed using complete peristomes from morphologically similar specimens.

## Filter complete specimens

To assess the accuracy of this reconstruction, totally complete peristomes are selected:

```{r interspec_filter}
data_sym <- symmetrize_sides(reordered_data, weighting = "exp")

#  Filter peristome without missing landmarks
filtered_landmarks_nonmerged <- filter_curves(data_sym, complete_only=TRUE)

# Merge R and L sides
data_complete <- merge_sides(filtered_landmarks_nonmerged)
```

::: {.callout-note collapse="true" title="dataset information"}
```{r TPS_rec_dataset}

table(filtered_landmarks_nonmerged$curve_info$Type) # count number of curves
cat("number of specimens included :",
    length(unique(filtered_landmarks_nonmerged$curve_info$Specimen))) # count number of specimens
```
:::

## Effect of missing landmark number

We started by studying the effect of missing landmark number of the reconstruction accuracy. Here, we deleted a random number of landmark, reconstruction those landmarks using *reconstruct_missingldm* function and computed the distance between original and reconstructed coordinates.

```{r nb_ldm_effect}
results_n_missing_ldm <- test_reconstruct_missingldm(data_complete, n_test = 30,
                                                     param_to_test = "n_missing_ldm",
                                                     verbose = FALSE)
```

The results, i.e. the error distances between original and reconstructed landmarks, standardized by whorl height, is summarized:

```{r nb_ldm_effect_res}
results <- results_n_missing_ldm$results
dist_by_ldm <- results_n_missing_ldm$dist_by_ldm

stats_quant <- results %>%
  group_by(Pourcentage) %>%
  summarise(
    Min  = min(Mean_dist_H, na.rm = TRUE),
    Q1   = quantile(Mean_dist_H, 0.25, na.rm = TRUE),
    Mean = mean(Mean_dist_H, na.rm = TRUE),
    Q3   = quantile(Mean_dist_H, 0.75, na.rm = TRUE),
    Max  = max(Mean_dist_H, na.rm = TRUE),
    .groups = "drop"
  )
plot(stats_quant$Pourcentage, stats_quant$Mean)
```

And then, we modelized the relationship between the mean distance by landmarks percent and the percent of missing landmarks using linear and exponential models. We compared model fitting using AIC values. and plot the best fitting model onto the data.

```{r nb_ldm_relationship}
# Delete the outlier
outlier <- which.max(stats_quant$Mean)
stats_quant <- stats_quant[-outlier, ]

# linear model ----
model_lin <- lm(Mean ~ Pourcentage, data = stats_quant) # linear model
summary(model_lin)

# Exponential model  ----

# Starting parameters
a_start <- min(stats_quant$Mean) * 0.9
b_start <- max(stats_quant$Mean) - a_start
c_start <- coef(lm(log(stats_quant$Mean - a_start + 1e-6) ~ stats_quant$Pourcentage))[2]

# exponential model y = a + b * exp(c * x)
model_exp <- nls(Mean ~ a + b * exp(c * Pourcentage),
                 data = stats_quant,
                 start = list(a = a_start, b = b_start, c = c_start),
                 control = nls.control(maxiter = 500, warnOnly = TRUE),
                 algorithm = "port")
summary(model_exp)

# Model comparison  ----
AIC(model_lin, model_exp)
```

➡️*The exponential model is the best fitting model (lower AIC).*

Here is the code used for the figure 9A in the manuscript):

```{r nb_ldm_relationship_plot}
# Predicted values of exponential model
stats_quant$Pred <- predict(model_exp)

# Plot data + exponential model
p <- ggplot(stats_quant, aes(x = Pourcentage)) +
  # geom_ribbon(aes(ymin = Min, ymax = Max), fill = "steelblue", alpha = 0.15) +
  # geom_ribbon(aes(ymin = Q1, ymax = Q3), fill = "steelblue", alpha = 0.35) +
  geom_point(aes(y = Mean), color = "black", size = 1) +
  geom_line(aes(y = Pred), color = "red", size = 1, linetype = "solid") +
  labs(
    title = "Effect of the proportion of missing landmarks on TPS-based interperistome\n reconstruction error.",
    x = "% missing landmarks",
    y = "Reconstruction error"
  ) +
  theme_minimal()
p
#ggsave("Effect_n_missing_ldm.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
```

Here, we added the standard deviation:

```{r nb_ldm_effect_ecartype}
ggplot(stats_quant, aes(x = Pourcentage)) +
  geom_ribbon(aes(ymin = Min, ymax = Max), fill = "steelblue", alpha = 0.15) +
  geom_ribbon(aes(ymin = Q1, ymax = Q3), fill = "steelblue", alpha = 0.35) +
  geom_point(aes(y = Mean), color = "black", size = 1) +
  labs(
    title = "Supplementary figure: Mean error distance per peristome vs % of missing landmarks",
    x = "% missing landmarks",
    y = "Reconstruction error"
  ) +
  theme_minimal()
```

➡️*The range of reconstruction error increase with the percent of missing landmarks.*

## 🆕 Impact donor number *(supp. info.)*

We started by deleting 20% of landmarks, either on the ventral or dorsal parts. This landmarks are then reconstructed using one to ten 'donors' complete specimens.

In this test, 20% of landmarks are deleted, a gaussian weighting function and no hierarchical structure are used.

```{r donor_nb}
results_n_donors <- test_reconstruct_missingldm(data_complete,
                                                param_to_test = "n_donors")

ggplot(results_n_donors$results, aes(x = as.factor(nb_donors), y = Mean_dist_H)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  #geom_jitter(width = 0.1, alpha = 0.4, color = "darkblue") +
  labs(
    x = "Number of donors",
    y = "Reconstruction distance error",
    title = "Supplementary figure: Reconstruction error distance vs number of donors"
  ) +
  theme_minimal()

# Normality test
by(results_n_donors$results$Mean_dist_H, results_n_donors$results$nb_donors, shapiro.test)

# Kruskal test
kruskal.test(Mean_dist_H ~ nb_donors, data = results_n_donors$results)
```

➡️*Changing donor number do not significantly change the reconstruction accuracy. However, range of error values is higher for low number of donors (1 and 2 donors). A low number of donors decreases function running time. Therefore, a default value of 5 donors was set.*

## 🆕 Test weighting functions *(supp. info.)*

We test the impact of weigthing functions choice on reconstruction accuracy. In this test, 20% of landmarks are deleted a maximal number of donors of 5 and no hierarchical structure are used.

```{r weight}
results_weighting_function <- test_reconstruct_missingldm(data_complete,
                               param_to_test = "weighting_function")

boxplot(Mean_dist_H ~ weighting_function, data = results_weighting_function$results,
        main = "Reconstruction error by weighting function",
        ylab = "Mean distance", xlab = "Weighting function",
        col = c("#66c2a5", "#fc8d62"))

# Shapiro test
shapiro.test(results_weighting_function$results$Mean_dist_H[results_weighting_function$results$weighting_function == "gaussian"])
shapiro.test(results_weighting_function$results$Mean_dist_H[results_weighting_function$results$weighting_function == "inverse"])
# Normality non respected

# Wilcoxon test
wilcox.test(Mean_dist ~ weighting_function, data = results_weighting_function$results)
```

➡️ *No weighting function significantly changes the reconstruction accuracy. Gaussian function was set by default.*

## 🆕 Test Hierarchy *(supp. info.)*

Impact of hierarchical structure is tested using four different structures :

-   No hierarchy, i.e. specimens are selected using only morphological proximity

-   Same specimen *→* All other peristomes

-   Same specimen *→* Same species *→* All other peristomes

-   Same specimen *→* Same species *→* Same genus *→* All other peristomes

In this test, 20% of landmarks are deleted, a gaussian weighting and a maximal number of donors of 3 are used.

```{r hierarchy}
results_Hierarchy <- test_reconstruct_missingldm(data_complete,
                                param_to_test = "Hierarchy")

results_Hierarchy$results$hierarchy <- recode(results_Hierarchy$results$hierarchy,
                                         "1" = "None",
                                         "2" = "Specimen",
                                         "3" = "Specimen -> Species",
                                         "4" = "Specimen -> Species -> Genus")

boxplot(Mean_dist_H ~ hierarchy, data = results_Hierarchy$results,
        main = "Reconstruction error by Hierarchical structure",
        ylab = "Mean distance", xlab = "Hierarchical structure",
        col = c(rainbow(4, start=.7, end=.1)))

# Normality test
by(results_Hierarchy$results$Mean_dist_H, results_Hierarchy$results$hierarchy, shapiro.test)

# Kruskal test
kruskal.test(Mean_dist_H ~ as.factor(hierarchy), data = results_Hierarchy$results)
```

➡️*No hierarchical structure significantly changes the reconstruction accuracy. This result is not surprising because phylogenetically close specimens are very often morphologically close.*

## Shape-related effect ?

Here is the code used for the [**figure 9B**]{.underline} in the manuscript:

```{r shape_effect_rec, message = FALSE, warning=FALSE, results='hide'}
set.seed(123)
results_Hierarchy <- test_reconstruct_missingldm(data_complete,
                                                 param_to_test = "none")
gpa <- Proc_peristome(data_complete,
                      slide = TRUE,
                      fixed_ldm = c(1, 25, 49))
pca <- gm.prcomp(gpa$coords)
pca_scores <- as.data.frame(pca$x[, ])

pca_scores$Mean_Distance_H <- results_Hierarchy$results$Mean_dist_H

# Percentage of variance explained
var_exp <- (pca$sdev^2) / sum(pca$sdev^2) * 100
pc1_lab <- paste0("PC1 (", round(var_exp[1], 1), "%)")
pc2_lab <- paste0("PC2 (", round(var_exp[2], 1), "%)")
```

```{r shape_effect_rec3, message = FALSE, warning=FALSE}
## Morphospace
p <- ggplot(pca_scores, aes(x = Comp1, y = Comp2, color = Mean_Distance_H)) +
  geom_point(size = pca_scores$Mean_Distance_H*15, shape = 16, stroke = 0, alpha = 0.8) +
  scale_color_viridis_c() +
  theme_minimal(base_size = 14) +
  labs(
    title = "PCA morphospace with peristomes colour-coded according \nto TPS-based reconstruction error.
",
    x = pc1_lab,
    y = pc2_lab,
    color = "Mean Distance / H"
  )
p
# ggsave("Shape_effect_TPS.pdf", plot = p, width = 8, height = 6, units = "in", device = "pdf")
```

::: {.callout-note collapse="true" title="Session info"}
```{r sessioninfo}
sessionInfo()
```
:::

# Summary

This vignette provides fully reproducible code to evaluate the main sources of methodological bias addressed in **AmmoniTools**. It complements the associated manuscript by documenting the computational workflow used to generate the bias assessment figures and supports transparency and reproducibility of the analyses.
